Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-04 p2/axioms.tex
Unified syntax

\begin{mylem}\mlabel{incomparable_is_independent}
Two different commands that are not assertion commands are independent 
iff the nodes they change are incomparable. In other words,
\[ \cxynv\indep\czwmv \Longleftrightarrow n\unrel m \]
if $\cxynv\neq\czwmv$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
The right-to-left part is a restatement of
\cref{ax_separate_commute,ax_separate_nobreaks}.

>>>>

\begin{mylem}\mlabel{incomparable_is_independent}
Two $\cxynv$ and $\czwmv$ are different commands,
and none of them is an assertion command, then
$\cxynv\indep\czwmv$ if and only if $n\unrel m$.
\end{mylem}
\begin{proof}
The right-to-left part is a restatement of
\cref{ax_separate_commute,ax_separate_nobreaks}.

### 2018-07-04 p2/def-commands.tex
Unified syntax

\begin{mydef}[Filesystem commands]
Filesystem commands are partial endofunctions on filesystems:
\[ \alpha: \setfs \nrightarrow \setfs \]
A command $\alpha$ is not defined on filesystems where it returns an error.
The commands are represented using triplets of the form
\[ \alpha = \caaaa{x}{y}{n}, \]
where $x,y\in\setv$ and $n\in\setn$.
The equivalence class $[x]$ is the \emph{input type} of the function;
$y$ is its \emph{output value} implicitly specifying the \emph{output type} $[y]$,
and $n$ is the node the command is applied to.
\end{mydef}
For example, $\cbfa{n}$ represents \textit{create}$(n,\valf)$,
and $\cdba{n}$ represents \textit{rmdir}$(n)$.

>>>>

\begin{mydef}[Filesystem commands]
Filesystem commands are partial endofunctions on filesystems,
which are not defined on filesystems where they return an error.
The commands are represented using triplets of the form
\[ \alpha = \caaaa{x}{y}{n}, \]
where $x,y\in\setv$ and $n\in\setn$.
The equivalence class $[x]$ is the \emph{input type} of the function;
$y$ is its \emph{output value} implicitly specifying the \emph{output type} $[y]$,
and $n$ is the node the command is applied to.
\end{mydef}
For example, $\cbfa{n}$ represents \textit{create}$(n,\valf)$,
and $\cdba{n}$ represents \textit{rmdir}$(n)$.

### 2018-07-04 p2/rec-complete.tex
More TODO
Unified syntax

% Reconciliation is maximal
% -------------------------
% This is where we're using |D|=1

@@

\begin{myth}\mlabel{rec_is_complete}
If $A$ and $B$ are simple sequences,
$S'\in\orderset{\ambnp}$,
and $S$ is a subsequence of $S'$ that
contains a command $\cxynv$ for which $\cxynv\nindep\bmanp$,
then $B\cc S$ either breaks all filesystems $A$ and $B$ are defined on
(that is, $S$ breaks all possible $\FS_B$ replicas),
or $S$ changes a node that $B$ has already changed to a different value,
that is, it overrides a change in $\FS_B$.
\end{myth}
Such an override could occur if a given node was modified differently in
$A$ and $B$ (to different values but to values of the same type), which 
our algorithm must treat as a conflict to be resolved by
by the user or a different system.
\begin{proof}
The proof is similar to the ones we have seen above.
Let $T$ be an arbitrary sequence in $\orderset{\bmanp}$.
Without loss of generality, we assume that $\cxynv$ is the first command in $S$
that is not independent of $T$.
If so, we can split $S$ into $S_0\cc\cxynv\cc S_1$ where $S_0\indep T$.

@@

From the proof it is apparent that this result depends on $|\setd|=1$.

>>>>

% Reconciliation is maximal
% -------------------------

@@

\begin{myth}\mlabel{rec_is_complete}
If $A$ and $B$ are simple sequences,
and $S$ is a subsequence of $S'\in\orderset{\ambnp}$ that
contains a command $\cxynv$ for which $\cxynv\nindep\bmanp$,
then $B\cc S$ either breaks all filesystems $A$ and $B$ are defined on
(that is, $S$ breaks all possible $\FS_B$ replicas),
or $S$ changes a node that $B$ has already changed to a different value,
that is, it overrides a change in $\FS_B$.
\end{myth}
Such an override could occur if a given node was modified differently in
$A$ and $B$ (to different values but to values of the same type), which 
our algorithm must treat as a conflict to be resolved by
by the user or a different system.
\begin{proof}
The proof is similar to the ones we have seen above.
Let $T$ be an arbitrary sequence from $\orderset{\bmanp}$.
Without loss of generality, we assume that $\cxynv$ is the first command in $S$
that is not independent of $T$.
If so, we can split $S$ into $S_0\cc\cxynv\cc S_1$ where $S_0\indep T$.

@@

% TODO Used elsewhere too!
From the proof it is apparent that this result depends on $|\setd|=1$.

### 2018-07-04 p2/rec-sound.tex
Unified syntax

To restate some results above, we therefore already know the following:
\begin{itemize}
\item $S=\amb$ and $T=\bma$ are simple as $A$ and $B$ are simple
\item $S\cap T = \amb \cap \bma = \emptyset$
\item $S^*$ is the largest subset of $S$ for which $S^*\indep T$, as
$\reca$ is the largest subset of $\amb$ for which $\reca\indep\bma$.
\end{itemize}
Our theorem is therefore proven if we prove the following lemma.
\newcommand{\condSimple}{(c1)}
\newcommand{\condDisj}{(c2)}
\newcommand{\condApr}{(c3)}
\begin{mylem}\mlabel{reconciliation_correct_part}
If
{\condSimple} $S$ and $T$ are simple sequences,
{\condDisj} $S\cap T=\emptyset$,
and {\condApr} $S^*$ is the largest subset of $S$ where $S^*\indep T$,
then
$\worksmbx{S,T}{S^*}$.
\end{mylem}

\begin{proof}
This is trivial unless $\worksmnilb{S,T}$, so we assume that it is the case.

The proof is similar to that of \cref{can_move_intersection}.
We mark all commands in $S$ that are in its subset $S^*$, and
we prove that $S$ can be transformed into $S'$
where all marked commands are at the beginning.
If so, then 
$\worksm{S}{S^*}$ based on \cref{worksextpostfix},
from which we get $\worksmbx{S,T}{S^*}$.

@@

Let therefore the marked command in $S$ be $\cxynv$
and the preceding unmarked command be $\czwmv$.
As $S$ is simple and $\wrks{S}$, from 
\cref{ax_distantrel_breaks,simple_distant_pairs}
we know that these commands can only be on incomparable nodes or form a construction or destruction pair.
In the first case, swapping the commands results in a sequence equivalent to $S$,
and we show that the other two cases are not possible as they would lead to a contradiction.

In the last two cases, we know that either $m=\parent(n)$ or $n=\parent(m)$.
We also know that because of {\condApr} there must be 
a command $\cqrov$ in $T$ which is not independent of $\czwmv$
as $\czwmv$ is not part of $S^*$.
\begin{gather*}
S = \cdots\cc  \czwmv\cc  \cxynv\cc  \cdots \\
T = \cdots\cc  \cqrov\cc \cdots
\end{gather*}
Based on \cref{incomparable_is_independent} we know this means that
either $m\descendantEq o$ or $o\descendantEq m$.
From {\condApr} we also know that $\cxynv\indep \cqrov$,
and so because of \cref{incomparable_is_independent},
$n\unrel o$.
We also know none of these commands is an assertion command, and 
from {\condDisj} that $\cqrov\neq\czwmv$.

>>>>

To restate some results above, we therefore already know the following:
\begin{itemize}
\item $S=\amb$ and $T=\bma$ are simple as $A$ and $B$ are simple
\item $S\cap T = \amb \cap \bma = \emptyset$
\item $S^*$ is the largest subset of $S$ for which $S^*\indep T$, as
$\reca$ is the largest subset of $\amb$ for which $\reca\indep\bma$.
\end{itemize}
Our theorem is therefore proven if we prove the following lemma.
\newcommand{\condSimple}{(c1)}
\newcommand{\condDisj}{(c2)}
\newcommand{\condApr}{(c3)}
\begin{mylem}\mlabel{reconciliation_correct_part}
If
{\rm\condSimple} $S$ and $T$ are simple sequences,
{\rm\condDisj} $S\cap T=\emptyset$,
and {\rm\condApr} $S^*$ is the largest subset of $S$ where $S^*\indep T$,
then
$\worksmbx{S,T}{S^*}$.
\end{mylem}
\begin{proof}
This is trivial unless $\worksmnilb{S,T}$, so we assume that it is the case.
The proof is similar to that of \cref{can_move_intersection}.
We mark all commands in $S$ that are in its subset $S^*$, and
we prove that $S$ can be transformed into $S'$
where all marked commands are at the beginning.
If so, then 
$\worksm{S}{S^*}$ based on \cref{worksextpostfix},
from which we get $\worksmbx{S,T}{S^*}$.

@@

Let therefore the marked command in $S$ be $\cxynv$
and the preceding unmarked command be $\czwmv$.
As $S$ is simple and $\wrks{S}$, from 
\cref{ax_distantrel_breaks,simple_distant_pairs}
we know that these commands can only be on incomparable nodes or form a construction or destruction pair.
In the first case, swapping the commands results in a sequence equivalent to $S$,
and we show that the other two cases are not possible as they would lead to contradiction.

In the last two cases, we know that either $m=\parent(n)$ or $n=\parent(m)$.
We also know that because of {\condApr} there must be 
a command $\cqrov$ in $T$ which is not independent of $\czwmv$
as $\czwmv$ is not part of $S^*$.
% \begin{gather*}
% S = \cdots\cc  \czwmv\cc  \cxynv\cc  \cdots \\
% T = \cdots\cc  \cqrov\cc \cdots
% \end{gather*}
Based on \cref{incomparable_is_independent} we know this means that
either $m\descendantEq o$ or $o\descendantEq m$.
From {\condApr} we also know that $\cxynv\indep \cqrov$,
and so because of \cref{incomparable_is_independent},
$n\unrel o$.
We also know none of these commands is an assertion command, and 
from {\condDisj} that $\cqrov\neq\czwmv$.
